---
title: "fireNASA Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{fireNASA-examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(fireNASA)
```

## FireNASA Package
## An API wrapper that accesses and extends the NASA Fireball dataset 

The fireNASA wrapper allows you to utilize NASA's freely available CNEOS Fireball dataset to investigate into the occurrence of fireballs by country, and by date. Fireballs or bolides are extremely bright meteors, which, when they hit the atmosphere, can burn as bright as or brighter than Venus at dawn, often becoming brighter than any object in the sky, other than the sun or moon. They last 5-10 seconds.  They occur much more often than they are seen, due to their short duration and interfering sky conditions (the sun or clouds). 

Note that this dataset isn't exhaustive. It is comprised of observations from US Government sensors and only includes the brightest fireballs. 

FireNASA contains one user-accessed function: `fireball_data()`, which   
- accesses the NASA Fireball dataset in R   
- automatically splits the date:time information into separate columns   
- provides the country name that the fireball was over when the observation was made   

With `fireball_data()`, you can choose to:   
- retrieve the entire NASA Fireball dataset as a dataframe in R     
- retrieve only a subset of the dataset, based on date/time, velocity, and/or radiated energy criteria   

Below are some examples of how to use `fireball_data()`

If you're interested, NASA's Fireball website can be accessed [here](https://cneos.jpl.nasa.gov/fireballs/intro.html))

***




## Retrieve the data
First, determine if you would like the entire dataset or a subset.  If you would like the entire dataset, then no parameters are required, and you can run the code below:

```{r message=FALSE}
df <- fireball_data()
```

```{r echo=FALSE}
knitr::kable(head(df), format="html")
```
We've shown the first six rows above. 

Note that `fireball_data()` does not retrieve observations that don't have latitude/longitude coordinates.


### Select fireball brightness   
Next let's try one of the filtering critera: peak brightness of the fireball (measured in x10^10 joules). Adding 'min_energy=15' as a parameter will limit your results to  fireballs that have an estimated total radiated energy greater than 15x10^10 joules.   
```{r message=FALSE}
df2 <- fireball_data(min_energy=15)

```


```{r echo=FALSE}
knitr::kable(head(df2), format="html")
```

***



## Explore the data


### Date and time separated   
`fireball_data()`, splits the date and time into seperate columns, making it easy to quickly create time series graphs that allow you to visualize fluctuations in fireball frequency over time.  Experiment with our code, below:

```{r results='hold', fig.height = 4, fig.width = 6, fig.align = "center"}
library(ggplot2)
library(dplyr)
ggplot(df %>% group_by(as.numeric(substring(date,1,4))) %>% dplyr::tally() %>% 
         rename(Year = `as.numeric(substring(date, 1, 4))`, No_ofFireballs = n),
       aes(Year, No_ofFireballs)) + 
  geom_line(color = "red", size = 1.5, linetype = 8)
```


### Country data added   
`fireball_data()` also adds a column called "Country", based on [Natural Earth's](https://www.naturalearthdata.com/) categroization (public domain). 

Note that all observations occuring over an ocean are given the label "OCEAN" and observations with bad coordinate data are labeled, "UNKNOWN".


### Graph countries by fireball observation count   
Below, we provide code that creates a bar chart of countries and the number of fireball events they have had.  Our code below uses `ggplot2`, but you can also do it quite quickly using the `dplry` package:

```{r results='hold', fig.height = 4, fig.width = 6, fig.align = "center"}
library(ggplot2)

#use run length encoding to get counts of all countries (excluding the countries "OCEAN" and "UNKNOWN")
country_cnts <- rle(sort(df$country[(df$country!="OCEAN"&df$country!="UNKNOWN")])) 
#store results in a dataframe
top5_df <- data.frame("country"=country_cnts$values,"count"=country_cnts$lengths, 
                      stringsAsFactors = FALSE)
#sort the dataframe by fireball count
top5_df <- top5_df[order(top5_df$count, decreasing = TRUE),]
#Take only the top 5
top5_df <- top5_df[0:5,]
print(top5_df)

#Now make a bar chart

ggplot() + geom_bar(aes(x=factor(country, levels=top5_df$country), 
                        y=count,fill=factor(country, levels=top5_df$country)),
                    data=top5_df,stat="identity") + 
  scale_fill_brewer(palette="Reds", direction=-1) +
  theme_dark() + labs(title = "Top 5 Countries", subtitle = "NASA fireball observations", 
                      fill="Country") + xlab("Country") + 
  theme(axis.text.x = element_text(angle = 60, hjust = 1))  

```


### Plot the fireball observations on a world map   
Since the dataset that `fireball_data()` retrieves contains location information you can plot this data on a world map and even highlight the top five countries that we found in the previous example. 

Note that this dataframe specifically includes two processed coordinate fields `lat_signed` and `lon_signed`. These are latitude and longitude in decimal degrees. Southern latitudes and Western longitudes have a negative value applied. This makes it very straightforward to plot on a map.

Our example uses `ggplot2` again as well as the `rnaturalearth` package (and associated data package) to provide map information and the `sf` (simple features) package as well as one of its helper packages (`rgeos`) to handle the map data for plotting.

```{r fig.height = 4, fig.width = 6, fig.align = "center", message=FALSE}

library(sf)
library(rgeos)
library(rnaturalearth)
library(rnaturalearthdata)


worldmap=ne_countries(returnclass="sf")
top5_worldmap = worldmap[(worldmap$name_long %in% top5_df$country),]


ggplot() + geom_sf(data=worldmap, aes(fill = NULL)) + 
  geom_sf(data=top5_worldmap, colour="red") + 
  geom_point(aes(x=df$lon_signed, y=df$lat_signed)) + 
  labs(title = "NASA Fireball Observations", 
       subtitle = "five countries with most observations outlined") +
  xlab("Longitude")+ ylab("Longitude")

```


