---
title: "plot-fireballs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{plot-fireballs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(fireNASA)
```

# FireNASA package 
## An API wrapper that accesses and extends the NASA Fireball dataset 

The fireNASA wrapper allows you to utilize NASA's freely available CNEOS Fireball dataset to investigate into the occurrence of fireballs by country, and by date. Fireballs or bolides are extremely bright meteors, which, when they hit the atmosphere, can as bright as or brighter than Venus at dawn.  They can sometimes be brighter than any object in the sky, except the sun or moon. They last 5-10 seconds.  They occur much more often than they are seen, partly due to their short duration and partly due to sky conditions (i. e. sun or clouds). 

Note that this dataset isn't exhaustive. It is comprised of observations from US Government sensors and only includes the brightest fireballs. 

FireNASA contains one user-accessed function: `fireball_data()` 

With the `fireball_data()` function you can:

- retrieve the entire NASA Fireball dataset as a dataframe in R   
- retrieve only a subset of the full NASA Fireball dataset based on date/time, velocity, and radiated energy criteria   

In addition to providing access to the NASA Fireball dataset in R, `fireball_data()` also automatically splits the date:time information into separate columns and provides the country name that the fireball was over when the observation was made. 


NASA's Fireball website can be accessed [here](https://cneos.jpl.nasa.gov/fireballs/intro.html))

***




## Retrieve the data
First, determine if you would like the entire dataset or a subset.  If you would like the entire dataset, then no parameters are required. 

```{r message=FALSE}
df <- fireball_data()
```

```{r echo=FALSE}
knitr::kable(head(df), format="html")
```
We've shown the first six rows above. Note that `fireball_data()` does not retrieve observations that don't have latitude/longitude coordinates, even if you request the entire dataset.


### Select fireball brightness   
Next let's try one of the filtering critera: peak brightness of the fireball (measured in 10^10 joules). Adding 'min_energy=15' delivers only fireballs with an estimated total radiated energy greater than 15x10^10 joules.   
```{r message=FALSE}
df2 <- fireball_data(min_energy=15)

```
```{r echo=FALSE}
knitr::kable(head(df2), format="html")
```

***



## Explore the data


### Separated time and date   
When you retrieve data using fireball_data(), the date and time information is automatically split into seperate columns in the dataframe. This makes it easy to quickly create a time series graph so you can visualize fluctuations in fireball frequency over time.

```{r results='hold', fig.height = 4, fig.width = 6, fig.align = "center"}
library(ggplot2)
library(dplyr)
ggplot(df %>% group_by(as.numeric(substring(date,1,4))) %>% dplyr::tally() %>% 
         rename(Year = `as.numeric(substring(date, 1, 4))`, No_ofFireballs = n),
       aes(Year, No_ofFireballs)) + 
  geom_line(color = "red", size = 1.5, linetype = 8)
```


### Working with Country Data   
When it retrieves the NASA data for you, `fireball_data()` adds a column for country. The column is based on [Natural Earth](https://www.naturalearthdata.com/) (public domain). Note that all observations occuring over an ocean are given the label "OCEAN". If any observations had bad coordinate data, you will see them labeled as "UNKNOWN".


### How to graph countries with most fireball observations   
You can use use the resulting dataframe to graph the top 5 countries for number of fireball events. This aggregation can be done quite quickly using the dplyr package but here we'll work with the dataframe that `fireball_data()` returned by just using base R functions and then plotting using `ggplot2`:

```{r results='hold', fig.height = 4, fig.width = 6, fig.align = "center"}
library(ggplot2)

#use run length encoding to get counts of all countries (excluding the countries "OCEAN" and "UNKNOWN")
country_cnts <- rle(sort(df$country[(df$country!="OCEAN"&df$country!="UNKNOWN")])) 
#store results in a dataframe
top5_df <- data.frame("country"=country_cnts$values,"count"=country_cnts$lengths, 
                      stringsAsFactors = FALSE)
#sort the dataframe by fireball count
top5_df <- top5_df[order(top5_df$count, decreasing = TRUE),]
#Take only the top 5
top5_df <- top5_df[0:5,]
print(top5_df)

#Now make a bar chart

ggplot() + geom_bar(aes(x=factor(country, levels=top5_df$country), 
                        y=count,fill=factor(country, levels=top5_df$country)),
                    data=top5_df,stat="identity") + 
  scale_fill_brewer(palette="Reds", direction=-1) +
  theme_dark() + labs(title = "Top 5 Countries", subtitle = "NASA fireball observations", 
                      fill="Country") + xlab("Country") + 
  theme(axis.text.x = element_text(angle = 60, hjust = 1))  

```


### Plot the fireball observations on a world map   
Since the dataset that `fireball_data()` retrieves contains location information we can plot this data on a world map and highlight the top five countries that we found in the previous example. 

Note that this dataframe specifically includes two processed coordinate fields `lat_signed` and `lon_signed`. These are latitude and longitude in decimal degrees. Southern latitudes and Western longitudes have a negative value applied. This makes it very straightforward to plot on a map.

This example uses `ggplot2` again as well as the `rnaturalearth` package (and associated data package) to provide map information and the `sf` (simple features) package as well as one of its helper packages (`rgeos`) to handle the map data for plotting.

```{r fig.height = 4, fig.width = 6, fig.align = "center", message=FALSE}

library(sf)
library(rgeos)
library(rnaturalearth)
library(rnaturalearthdata)


worldmap=ne_countries(returnclass="sf")
top5_worldmap = worldmap[(worldmap$name_long %in% top5_df$country),]


ggplot() + geom_sf(data=worldmap, aes(fill = NULL)) + 
  geom_sf(data=top5_worldmap, colour="red") + 
  geom_point(aes(x=df$lon_signed, y=df$lat_signed)) + 
  labs(title = "NASA Fireball Observations", 
       subtitle = "five countries with most observations outlined") +
  xlab("Longitude")+ ylab("Longitude")

```


